---
title: "Factor analysis of Hopkins Symptom Checklist 25 (English) in HIV-infected individuals who are not first-language English speakers"
author: 'Peter Kamerman, Prinisha Pillay, Antonia Wadley' 
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  github_document:
    html_preview: false
---

## Load required packages and set chunk options
```{r miscellaneous.1, include = FALSE}
# Load libraries
library(pander)
library(readr)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(GPArotation) # for fa function in 'psych'

# Set seed
set.seed(123)

# Set knitr chunk options
opts_chunk$set(echo = TRUE,
               warning = FALSE,
               message = FALSE,
               cache = FALSE,
               dev = c('png'),
               tidy = FALSE)
```

### Import data
```{r import.1}
# Read csv
data <- read_csv('./data/factor-analysis.csv')
```

### Inspect
```{r inspect.1, include = FALSE}
# Inspect the data
head(data)
tail(data)
str(data)
```

### Clean
```{r clean.1, echo = TRUE}
# Convert ID to factor
data_clean <- data %>%
    select(-pain) %>%
    mutate(id = factor(id))
```

### Internal consistency
_**Use polychoric correlation matrix because data are ordinal integers, and therefore Pearson's correlation matrix does not apply**_
```{r ic.1}
# Remove ID column
data_ic <- data_clean[ , -1]

# Anxiety subset
anx_ic <- data_ic %>%
    select(starts_with('anx'))
## Get a summary of the data
kable(describe(anx_ic))
## Generate and view polychoric correlation matrix
anx_pc <- polychoric(anx_ic)
## Generate internal consistency coefficient alpha
anx_alpha <- alpha(anx_pc$rho)
### Summary stats
summary(anx_alpha)
### Drop-one analysis
#### r.cor = Item whole correlation corrected for 
#### item overlap and scale reliability.
#### r.drop = Item whole correlation for this item 
#### against the scale without this item
anx_alpha$item.stats

# Depression subset
dep_ic <- data_ic %>%
    select(starts_with('dep'))
## Get a summary of the data
kable(describe(dep_ic))
## Generate and view polychoric correlation matrix
dep_pc <- polychoric(dep_ic)
## Generate internal consistency coefficient alpha
dep_alpha <- alpha(dep_pc$rho)
### Summary stats
summary(dep_alpha)
### Drop-one analysis
#### r.cor = Item whole correlation corrected for 
#### item overlap and scale reliability.
#### r.drop = Item whole correlation for this item 
#### against the scale without this item
dep_alpha$item.stats

# Total questionnaire
## Get a summary of the data
kable(describe(data_ic))
## Generate and view polychoric correlation matrix
data_pc <- polychoric(data_ic)
## Generate internal consistency coefficient alpha
pander(summary(alpha(data_pc$rho)), 
       style='rmarkdown', 
       justify='left', 
       round = 2, 
       caption='HSCL25: summary data')
```

### Factor analysis - initial
_**Use polychoric correlation matrix because data are ordinal integers, and therefore Pearson's correlation matrix does not apply**_
```{r fa.1}
# Print parallel plots to estimate factor solution
fa.parallel(data_ic,
            fm='ml',
            cor='poly')

# Complete factor analysis - use maximum liklihood (ml) method
## Use nfactor = 4 based on fa.parallel plot
## Thereafter repeat analysis using 2 factors
## Test oblique rotation (Oblimin), assuming correlation between factors

### Oblimin: 4 factor
fa_4 <- fa.poly(data_ic,
                nfactors = 4,
                rotate ='oblimin',
                fm ='ml',
                cor = 'poly',
                global = F)
### Oblimin: 2 factor
fa_2 <- fa.poly(data_ic,
                nfactors = 2, 
                rotate ='oblimin',
                fm = 'ml',
                cor = 'poly',
                global = F)

# Interpret
## Oblimin: 4 factor (suggested)
fa.diagram(fa_4,
           cut = 0,
           digits = 3)

print(fa_4, 
      cut = 0,
      digits = 3)

## Oblimin: 2 factor
fa.diagram(fa_2, 
           cut = 0,
           digits = 3)

print(fa_2,
      cut = 0,
      digits = 3)
```
## Session information
```{r SessionInfo}
sessionInfo()
```




